---
title: "Kevin Ortega"
output: html_document
date: "2024-09-27"
---


This assignment has three questions:  

1. in a class of 35 students, Dr. J bets $1 against student's $1  that at least 2 students have the same birthday, month, day.
Find the probability thatat least two student have the same birthday.  If 26 students in the class take the bet how much is Dr. J expected to win/lose.   How much should Dr. J wager against the student's $1 to make this a fair bet?

2. Assume class size is 26 instead of 35.  With the same bet ($1 ot $1) what is the probability Dr. J wins the bet now.  If 20 of the 26 students make the bet, what is Dr. J's expected profit/loss?

3. What size class (how many students) makes the $1 againt $1 a fair bet?

##Question 1

```{r setup, include=FALSE}
set.seed(123) # Set a seed for reproducibility
```

```{r}
# Define function to simulate the birthday problem
birthday_simulation <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  any(duplicated(birthdays)) # Check if there are duplicates
}

# Run the simulation 100 times and save probabilities in a vector
num_simulations <- 100
class_size <- 35
probabilities <- numeric(num_simulations)

for (i in 1:num_simulations) {
  # Run 1000 simulations to estimate the probability for each iteration
  num_duplicates <- replicate(1000, birthday_simulation(class_size))
  probabilities[i] <- mean(num_duplicates)
}

# Print the mean of the probabilities
mean_prob <- mean(probabilities)
cat("Mean =", mean_prob, "\n")

# Plot a histogram of the probabilities
hist(probabilities, main = "Histogram of Estimated Probabilities",
     xlab = "Probability", col = "skyblue", border = "white")

# Calculate a 95% confidence interval
conf_interval <- quantile(probabilities, probs = c(0.025, 0.975))
cat("95% Confidence Interval = [", conf_interval[1], ", ", conf_interval[2], "]\n")
```


```{r}
EV= mean_prob*26-(1-mean_prob)*26
print(paste("ev=",EV))
```

```{r}
FAIR=(mean_prob)/(1-mean_prob)
print(paste("fairbet=",FAIR))
```
Find the probability that at least two student have the same birthday= 81%
If 26 students in the class take the bet how much is Dr. J expected to win/lose. - He is expected to win.
How much should Dr. J wager against the student's $1 to make this a fair bet? - Dr.J would need to bet $4.46 dollars for it to be a fair bet.

##Question2
```{r}
# Define function to simulate the birthday problem
birthday_simulation <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  any(duplicated(birthdays)) # Check if there are duplicates
}

# Run the simulation 100 times and save probabilities in a vector
num_simulations <- 100
class_size <- 26
probabilities <- numeric(num_simulations)

for (i in 1:num_simulations) {
  # Run 1000 simulations to estimate the probability for each iteration
  num_duplicates <- replicate(1000, birthday_simulation(class_size))
  probabilities[i] <- mean(num_duplicates)
}

# Print the mean of the probabilities
mean_prob <- mean(probabilities)
cat("Mean =", mean_prob, "\n")

# Plot a histogram of the probabilities
hist(probabilities, main = "Histogram of Estimated Probabilities",
     xlab = "Probability", col = "skyblue", border = "white")

# Calculate a 95% confidence interval
conf_interval <- quantile(probabilities, probs = c(0.025, 0.975))
cat("95% Confidence Interval = [", conf_interval[1], ", ", conf_interval[2], "]\n")
```

```{r}
EV= mean_prob*20-(1-mean_prob)*20
print(paste("ev=",EV))
```

```{r}
FAIR=(mean_prob)/(1-mean_prob)
print(paste("fairbet=",FAIR))
```

Assume class size is 26 instead of 35.  With the same bet ($1 ot $1) what is the probability Dr. J wins the bet now.  = 
If 20 of the 26 students make the bet, what is Dr. J's expected profit/loss?

##Question 3

```{r}
# Define function to simulate the birthday problem
birthday_simulation <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  any(duplicated(birthdays)) # Check if there are duplicates
}

# Run the simulation 100 times and save probabilities in a vector
num_simulations <- 100
class_size <- 23
probabilities <- numeric(num_simulations)

for (i in 1:num_simulations) {
  # Run 1000 simulations to estimate the probability for each iteration
  num_duplicates <- replicate(1000, birthday_simulation(class_size))
  probabilities[i] <- mean(num_duplicates)
}

# Print the mean of the probabilities
mean_prob <- mean(probabilities)
cat("Mean =", mean_prob, "\n")

# Plot a histogram of the probabilities
hist(probabilities, main = "Histogram of Estimated Probabilities",
     xlab = "Probability", col = "skyblue", border = "white")

# Calculate a 95% confidence interval
conf_interval <- quantile(probabilities, probs = c(0.025, 0.975))
cat("95% Confidence Interval = [", conf_interval[1], ", ", conf_interval[2], "]\n")
```


```{r}
EV= mean_prob*20-(1-mean_prob)*20
print(paste("ev=",EV))
```


In order for it to be a fair bet the class size would need to be 23.